---
title: "Final Project - Yelp Recommender System"
author: "Juliann McEachern, Rajwant Mishra,Christina Valore"
date:  " July 16, 2019"
output: 
  html_document:
    theme: flatly
    highlight: pygments
    toc_depth: 2
    df_print: paged
    code_folding: hide
---

# Overview

Add overview here. 

```{r, echo=F}
## data processing packages
library(dplyr); library(tidyr);library(jsonlite)
##formatting packages
library(knitr); library(kableExtra); library(default)
##visualization packages
library(ggplot2)
##recommender packages
library(recommenderlab); library(Metrics); library(lsa); library(diveRsity)
```

### Data Aquisition

@ Raj - Add textual explanation of process. What data is available? Explain variables we have to work with. 

```{r include=FALSE}
# Read JSON files from repository 
# Business Data 
business<-stream_in(file("data/yelp_training_set_business.json"))

# user Data 
user <-stream_in(file("data/yelp_training_set_user.json"))

# checkin Data 
checkin <-stream_in(file("data/yelp_training_set_checkin.json"))

# review  Data 
review <-stream_in(file("data/yelp_training_set_review.json"))
```

### Data Transformations

@ Raj - Explain process. What was included/excluded? Why? Ie: 

We changed the business and user columns name for clarity to better identify our variables. 

```{r include=FALSE}
# Business Data

## create dataframe
df_business<- business[ , order(names(business))] %>% # sort alphabetically
  .[,c(3,1,2,7,9,10,12,13,11)]  # reorder columns keeping STARS for Training set at index 11

df_business <- mutate(df_business, cat_string = sapply(categories, toString)) # save categories as strings.

## rename headers 
HeaderName_bus <- c( "bus_city",
                  "business_id",
                  "bus_categories",
                  "bus_name",
                  "bus_open" ,
                  "bus_review_count",
                  "bus_state",
                  "bus_type",
                  "bus_stars",
                  "bus_categories_all")

names(df_business) <- HeaderName_bus

# User Data
# Votes is dataframe in the user dataset , we will read this data and then bind columns with user data.
df_user <- user %>% select( -votes) %>% bind_cols(.,user$votes) %>% .[,c(1,2,4,5,3,6,7,8)]
df_user_long <- gather(df_user,key= "mood", value = "rating",funny,useful,cool)

# Test User
# Updating the name of the Users Votes column to user_funny , user_usefull, user_cool
HeaderName_usr <- c("user_id","usr_name","usr_review_count", "usr_type","usr_average_stars","usr_funny","usr_useful","usr_cool")


names(df_user) <- HeaderName_usr

#checkins data

#         '0-0': (number of checkins from 00:00 to 01:00 on all Sundays),
#         '1-0': (number of checkins from 01:00 to 02:00 on all Sundays), 
#         ... 
#         '14-4': (number of checkins from 14:00 to 15:00 on all Thursdays),
#         ...
#         '23-6': (number of checkins from 23:00 to 00:00 on all Saturdays)

# checkin_info is a dataframe with the above structure of the data, we would extract this info and append it to the Checkin Table against the Business ID.
df_checkin <- checkin  %>% select( -checkin_info) %>% bind_cols(.,checkin$checkin_info) # %>% .[,c(1,2,4,5,3,6,7,8)]


#Reviews Data 
df_review_vote <- review$votes
df_review <- review  %>% select( -votes) %>% bind_cols(.,df_review_vote) 
HeaderName_rev <- c("user_id" ,
                "review_id",
                "rev_stars",
                "rev_date" ,
                "rev_text" ,
                "rev_type" ,
                "business_id",
                "rev_funny"   ,
                "rev_useful" ,
                "rev_cool")
names(df_review) <- HeaderName_rev

```

We then joined our dataframes using the `Business_ID` and `User_ID` as unique keys. 

```{r}
#----------------------------------------------------------------------------------------------
# Building Main data set by Joining all the above data sets on key of Business_ID and User_ID
#----------------------------------------------------------------------------------------------

#Group all reviews that have a business with a star rating
review_bus  <- inner_join(df_review,df_business,by=("business_id"))

#Group all reviews that have and User's data
review_user <- inner_join(df_review,df_user,by=("user_id"))

# joining all the "Review and Users" information with  the "Review and Business"  Dataset. 
# Moving Key in the Begning 
df_main <- inner_join(review_bus,review_user[,-c(2,3,4,5,6,8,9,10)], 
                              by =c("business_id","user_id")) %>% .[,c(1,2,7,3,4,5,6,8:26)]
# Appending the Checkins information for the Business 
df_main_chk <- inner_join(df_main,df_checkin[,-c(1)],by=("business_id"))

```

## Data Introduction  {.tabset .tabset-fade .tabset-pills}

### Business
This dataset list business information with average business rating, along with category of the Business.
<center>**Business Train Data** </center>
 `r head(df_main_train_business)`
 
### Review
Review is main dataset here, it holds link to Users and Business and the corresponding Text Reviews are listed along with ratings.

<center>**Review Train Data** </center>
`r head(df_main_train_review)`

### User
Users Holds all the information about user, average rating of the Users given so far, count of the review and different count of Votes on "Cool", "Useful" and "funny".

<center>**User Train Data** </center>
`r head(df_main_train_user) `

### Checkin 
Checkin information is round the clock and 7 days data of the Customer's checkin. Column name start with hour-Day 
here hours [00 to 23]- [0 to 6]

<center>**Checkin Train Data** </center>
`r head(df_main_train_checkin)`

## Final Data: 
* We renamed All the Columns from Business (bus_),User(usr_) and Review (rev_) dataset, so that we can identify them from the Big dataset
* Review dataset will be combined with User and Business dataset, with User_ID and Business_Id as key
* Final Dataset is then combined with Checkin information of the Businesses 


```{r echo=FALSE}
# DT::datatable(head(DF_MAIN,10), caption = "Review Data")
# head(DF_MAIN_CHK,10) %>% kable(caption="Data with Checkins Info") %>% kable_styling()

#head(df_main_chk,10) 


df_main %>% select(user_id) %>% group_by(user_id) %>% tally() %>% arrange(desc(n))
```

## Data Exploration {.tabset .tabset-fade .tabset-pills}

WIP

## Data Visualizations {.tabset .tabset-fade .tabset-pills}

WIP

## Data Preparation for Model {.tabset .tabset-fade .tabset-pills}

## Matrix Building 
We converted our raw ratings data into a user-item matrix to test and train our subsequent recommender system algorithms.

```{r}

# create user item matrix with 500 items only as I was getting storage error.
# https://github.com/tidyverse/tidyr/issues/426

df_main$user_id <- as.factor(df_main$user_id)
df_main$business_id <- as.factor(df_main$business_id)

mat <- df_main %>% 
  select(user_id, business_id, rev_stars) %>%
  dcast(user_id ~ business_id)




#rownames(ui_matrix_star)<-ui_matrix_star$user_id # set row names to user_id
#ui_matrix_star<-ui_matrix_star %>% select(-user_id) %>% as.matrix()# remove user_id from columns
#umat <- as(ui_matrix_star,"realRatingMatrix") # save real ratings for algo 
#real_ui_matrix_star <- as(ui_matrix_star,"realRatingMatrix") # save real ratings for algo 

# preview matrices 
#as.data.frame.array(ui_matrix_star) %>% head() %>% kable(caption="Preview of User-Item Matrix (User-Business)") %>% kable_styling() %>% scroll_box()



```


## Training and Test Subsets

Finally, our data was split into training and tests sets for model evaluation of both two recommender algorithms. We split our data with 10 k-folds using the `recommendaerlab` package. 80% of data was retained for training and 20% for testing purposes.

```{r train-test}
# evaluation method with 90% of data for train and 10% for test
set.seed(1000)

evalu <- evaluationScheme(real_ui_matrix_star, method="split", train=0.8, given=0)

# Prep data
ratings_train <- getData(evalu, 'train')# Training Dataset 
ratings_test_known <- getData(evalu, 'known') # Test data from evaluationScheme of type KNOWN
ratings_test_unknown <- getData(evalu, 'unknown') # Unknow datset used for RMSE / model evaluation

```



## Algorithm{.tabset .tabset-fade .tabset-pills}


## Conclusion {.tabset .tabset-fade .tabset-pills} 

## Accuracy Metrics {.tabset .tabset-fade .tabset-pills}

# References
* [**Tidyr Issue**](https://github.com/tidyverse/tidyr/issues/426)
* [**Data Overview**](https://www.kaggle.com/c/yelp-recsys-2013/overview)



```{r}

```

<hr><hr>
