---
title: "Final Project - Yelp Recommender System"
author: "Juliann McEachern, Rajwant Mishra,Christina Valore"
date:  " July 16, 2019"
output: 
  html_document:
    theme: sandstone
    highlight: pygments
    toc_depth: 2
    df_print: paged
    code_folding: hide
---
# Overview 

Our project used data from Kaggle's 2013 Yelp Challenge. This challenge included a subset of Yelp data from the metropolitan area of Phoenix, Arizona. Our data takes into account user reviews,ratings, and check-in data for a wide-range of businesses. 

```{r,echo=F,comment=F,message=F,warning=F,prompt=F}
## data processing packages
library(dplyr); library(tidyr);library(jsonlite);library(stringr)

##formatting packages
library(knitr); library(kableExtra); library(default); library(hrbrthemes)

##visualization packages
library(ggplot2)

##recommender packages
library(recommenderlab); library(Metrics); library(lsa); library(diveRsity)

# global options

## knit sizing
options(max.print="100"); opts_knit$set(width=75) 

## augment chunk output
opts_chunk$set(echo=T,cache=F, tidy=T,comment=F,message=T,warning=T) #change message/warning to F upon completion

## set table style for consistency
default(kable) <- list(format="html")
default(kable_styling)  <- list(bootstrap_options = "hover",full_width=T, font_size=10)
default(scroll_box) <- list(width = "100%")
```


## Data Aquisition

All the data is multiarray of Json, which means file is a collection of json data. We will use  `stream_in` function, which parses json data line-by-line from our files stored within the data folder of our repository. 

```{r, message=F, warning=F}
# Business Data 
business<-stream_in(file("data/yelp_training_set_business.json"),verbose = F)
# user Data 
user <-stream_in(file("data/yelp_training_set_user.json"),verbose = F)
# checkin Data 
checkin <-stream_in(file("data/yelp_training_set_checkin.json"),verbose = F)
# review  Data 
review <-stream_in(file("data/yelp_training_set_review.json"),verbose = F)
```


## Data Transformations {.tabset .tabset-fade .tabset-pills}

### Business

We choose to limit the scope to our recommender system to only businesses with tags related to food and beverages. There were originally 508 unique category tags listed within our business data. We manually filtered 112 targeted categories to subset our data. 

We applied additional transformation to remove unnessacary data. There were 1224 business in our data that were permanently closed. These companies accounted for 9.8% of all businesses, which were subsequently removed from our data. There were also 3 businesses in our dataset from outside of AZ that we allso removed. 

As a result of our transformations, our recommender data now consists of 4828 unique businesses. This data is previewed below: 

```{r, warning=F, comment=F}
# dropped open, neighborhoods(no data), full address, and review count, type (all business)
df_business<-business %>% filter(open == "TRUE", state=="AZ") %>% select(-open, -neighborhoods, -full_address, -review_count, -type, -stars) %>% mutate(categories = sapply(categories, toString)) 

# create category filter
filter <- 'Argentine| Burmese| Cambodian| Cocktail Bars|Laotian|Lebanese|Live/Raw Food|Russian|African|Champagne Bars|Kosher|Modern European|Scandinavian|Taiwanese|Tapas/Small Plates|Afghan|Brazilian|Food Trucks|Shaved Ice|Wineries|Dim Sum|Ethiopian|Fondue|Hookah Bars|Persian/Iranian|Peruvian| Polish| Seafood Markets| Tapas Bars|Halal|British| Cheese Shops|German|Spanish|Cheesesteaks|Cuban|Do-It-Yourself Food|Gastropubs|Salad|Creperies|Soup|Chocolatiers & Shops|Filipino|Food Stands|Fruits & Veggies|Meat Shops|Mongolian|Soul Food|Comfort Food| Irish|Fish & Chips|Cajun/Creole|Caribbean|Pakistani|Southern|Candy Stores|Vegan|Latin American|Breweries|French|Gay Bars|Korean|Gluten-Free|Hawaiian|Farmers Market|Vegetarian|Middle Eastern|Ethnic Food|Indian|Pubs|Chicken Wings|Dive Bars| Juice Bars & Smoothies|Vietnamese|Cafes|Wine Bars|Bagels|Diners|Hot Dogs|Tex-Mex|Donuts|Greek|Thai| Desserts|Mediterranean|Beer| Wine & Spirits|Seafood|Sushi Bars| Lounges|Steakhouses|Buffets|Japanese|Sports Bars|Delis|Bakeries|Specialty Food|Breakfast & Brunch|Ice Cream & Frozen Yogurt|Burgers|Italian| Chinese|Coffee & Tea|American (New)|Sandwiches|Fast Food|Pizza|American (Traditional)|Bars|Mexican|Food| Restaurants'

# filter businesses using filter & factor categories
df_business <- df_business %>% filter(str_detect(categories, filter)) %>% mutate(categories = as.factor(categories))
```


```{r, echo=F}
#preview business
df_business %>% head() %>% kable(caption="Preview Business Data") %>% kable_styling()
```

### Review

We subset our review data from the subset of food and beverage businesses. This dropped our review data from 229,907 to 165,823 reviews. 

```{r}
#subset reviews & remove columns (type)
df_review <- review %>% filter(business_id %in% df_business$business_id)%>% select(-type)

# assemble ratings data (funny/useful/cool) into singular columns.
df_review <- do.call(data.frame, df_review)

# remove factor from user_id
df_review$user_id <- as.character(df_review$user_id)
df_review$business_id <- as.character(df_review$business_id)
```

```{r, echo=F}
#preview reviews
df_review %>% select(-text) %>% head() %>% kable(caption="Preview Review Data (without Review Text)") %>% kable_styling()

#preview review text

df_review %>% select(text) %>% head(1) %>% kable(caption="Preview of a Singular Review Text") %>% kable_styling()
```


### User 

We applied a similar filter to users to subset our data based on only our selected businesses. This decreased our user data from 43,873 to 35,268 distinct user_id observations. 

The dataframe preview below shows aggregate user data for all reviews an individual user provided for yelp within our data selection. 

```{r, echo=F}
#subset user & remove columns (type)
df_user <- user %>% filter(user_id %in% df_review$user_id) %>% select(-type)

# assemble ratings data (funny/useful/cool) into singular columns.
df_user <- do.call(data.frame, df_user)
```

```{r, echo=F}
#preview users
df_user %>% head() %>% kable(caption="Preview User Data") %>% kable_styling()
```

### Checkins

Lastly, we evaluated our checkins data and applied our business filter once more. This decreased our checkin observations from 8282 to 4423. This change tells us that no checkin data is available for 8.4% of businesses in our subset. 

```{r, echo=F}
#subset checkin & remove columns (type)
df_checkin <- checkin %>% filter(business_id %in% df_business$business_id) %>% select(-type)

# unpack checkin_info into singular columns.
### there are 168 checkin column variables. wait to see how data is used before unpacking with do.call.
#df_checkin <- do.call(data.frame, df_user)
```

```{r, echo=F}
#preview users
df_checkin %>% head() 

### unable to apply kable style without unpacking checkin_info TBD
```

## Merge Data 

Next, we created our main dataframe business and review dataframes using the `Business_ID` and `User_ID` as unique keys. This dataframe will be referenced later on when building our recommender matrices and algorithms. 

```{r}
df_main <-df_business %>% inner_join(df_review, by="business_id")
```

```{r, echo=F}
#preview main df
df_main %>% head() 
```

## Visualize Data

**CLEAN UP SECTION**

The graphs below show geographic trends in our data as well as our most popular categories. 

```{r}
# All Business City
df_business %>% select (business_id, city) %>% group_by(city) %>% dplyr::summarise(Count=n()) %>%   ggplot(aes(x = city, y = Count)) +
         geom_bar(stat = "identity") +
          theme(axis.text.x=element_text(angle=90, hjust=1))+
    xlab(label = "Business City")+
  geom_text(aes(label = Count), colour = "blue", fontface = "bold" ,position = position_stack(vjust = 1.1),check_overlap=TRUE)


# All Business City by Focus Category

## add graph later
  
```

CONSIDER Including plot showing something like number of reviews for each business. Review date by month is pretty consistant and not telling of much on its own. 
```{r, echo=F}
library(lubridate)
# Avg Rating Given by User
df_review %>% select (stars,date) %>% 
  mutate(month = month(as_date(date))) %>% group_by(month) %>% 
  dplyr::summarise(Count=n()) %>%  
  ggplot(aes ( x= factor(month),y = Count)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x=element_text(angle=00, hjust=1))+
  xlab(label = "Avg Rating Given by User")


str_split(df_review$rev_date[1],"-",n=3,simplify = TRUE)[2]
```

The plot below shows the distribution of our average ratings given by all users. 
```{r, echo=F}
# Avg Rating Given by User
df_user %>% select (user_id,average_stars,review_count) %>%
  ggplot(aes(x = round(average_stars), y = review_count)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x=element_text(angle=00, hjust=1))+
  xlab(label = "Avg Rating Given by User")
```



## Matrix Building 

We converted our raw ratings data into a user-item matrix to test and train our subsequent recommender system algorithms. The matrix was saved as a realRatingMatrix for processing purposes later on using the `recommenderlab` package. 

```{r}
matrix_data <- df_main %>% select(user_id, business_id, stars) %>% mutate(user_id=as.factor(user_id), business_id=as.factor(business_id), stars=as.numeric(stars))

ui_mat <- as(matrix_data,"realRatingMatrix")
```

## Training and Test Subsets

Finally, our data was split into training and tests sets for model evaluation of both two recommender algorithms. We split our data with 10 k-folds using the `recommenderlab` package. 80% of data was retained for training and 20% for testing purposes.

```{r train-test}
# evaluation method with 80% of data for train and 20% for test
set.seed(1000)

evalu <- evaluationScheme(ui_mat, method="split", train=0.8, given=0, goodRating=1, k=10)

# Prep data
train <- getData(evalu, 'train')# Training Dataset 
dev_test <- getData(evalu, 'known') # Test data from evaluationScheme of type KNOWN
test <- getData(evalu, 'unknown') # Unknow datset used for RMSE / model evaluation
```




# Recommender Algorithm

## Algorithm 1 (Raj)
```{r}
# I am planning to see how I can use Business's rating and rev_funny,rev_useful,rev_cool
# and see how users are rating against these parameters, I would check cosine similarlty of user 
# rating with these info and recommend some similar  Business to the Users.
```

## Algorithm 2 (Christina)

## Algorithm 3 (Juliann)

## Analysis

Compare algorithms performance. Select most effective to build recommender system.

# Recommender

Test system

# Conclusion

Final conlusion. Explain limitations of system. Make recommendations for future improvements.  

# References
* [**Data Overview**](https://www.kaggle.com/c/yelp-recsys-2013/overview)

