---
title: "Project Proposal"
author: "Rajwant Mishra"
date: "July 5, 2019"
output:
  html_document:
    df_print: paged
    highlight: tango

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load data
library(tidyverse)
library(rjson)
library(XML)
library(xml2)
library(ggplot2)
library(dplyr)
library(maps)
library(ggmap)
library(mongolite)
library(lubridate)
library(gridExtra)
library(tidyverse)
library(httr)
library(jsonlite)
library(DT)
library(rjson)
library(readxl) 
#install.packages("DT")
library(DT)

```

```{r eval=FALSE, include=FALSE}
workDir <- getwd()

filePath = paste0(workDir,"/data")
g_max <- 1048576 
ctype = c("text","text","text","text","text","numeric","numeric","text","text","text","text","text","text","text")
Full_Data <-read_excel(path=paste0(filePath,"/","DATA612.xlsx"),sheet="Sheet1",guess_max =g_max)

head(Full_Data)
custA <- Full_Data[sample(nrow(Full_Data),1000),]
```

```{r eval=FALSE, include=FALSE}
head(custA)
```

```{r eval=FALSE, include=FALSE}
# [1] Planning Document Find an interesting dataset and describe the system you plan to build out. If you would like to use one of the datasets you have already worked with, you should add a unique element or incorporate additional data. (i.e. explicit features you scrape from another source, like image analysis on movie posters). The overall goal, however, will be to produce quality recommendations by extracting insights from a large dataset. You may do so using Spark, or another distributed computing method, OR by effectively applying one of the more advanced mathematical techniques we have covered. There is no preference for one over the other, as long as your recommender works! The planning document should be written up and published as a notebook on GitHub or in RPubs.Please submit the link in the Unit 4 folder, due Thursday, July 5.
# 
# [2] Presentation. Make a five-minute presentation of your system in our final meetup on Tuesday. If you're not able to attend the meetup, you're responsible for either recording your presentation, or scheduling one-on-one time to deliver your presentation prior to the meetup. You should be prepared to present on Tuesday. You should use this project to showcase some of the concepts that you have learned in this course, while delivering on the (probably) less familiar Spark platform. You are welcome to submit a compelling alternative proposal (subject to approval), such as implementing a recommender system using in Microsoft Azure ML Studio or with Google TensorFlow, or building out an application of a certain complexity using another tool. You may work in a small group (2-3) on this assignment.
# 
# [2] Implementation. In this final project deliverable, you'll build out the system that you describe in your planning document. This will be due on Thursday and must be turned in as an RMarkdown file or a Jupyter notebook, and posted to GitHub or RPubs.com.

```

### Introduction About Data {.tabset .tabset-fade .tabset-pills}
 
 This is the random data of the sales by customeracross all differnet categories of the products offered at Firm ABC Corp.
 
#### Data Fileds  
 |    Field NAME   |                 Description                |                   Note                   |
|:---------------:|:------------------------------------------:|:----------------------------------------:|
|     Customer    | Customer  More than one record in the data  | More than One record   for each Customer |
|      Brand      | Category Name of the   Products            | More than One record   for each Customer |
|    Order_Qty    | Order Quantity                             |                                          |
|    Orer_Cost    | $ Value of Order                           |                                          |
|       city      | City of the Customer                       |                                          |
|      state      | State of the Customer                      | Missing for few Entries                  |
|       Year      | Year of the Order                          |                                          |
|     Country     | Country                                    |                                          |


#### Sample


| Customer | Brand | Order_Qty | Orer_Cost |     city    | state | Year | Country |
|:--------:|:-----:|:---------:|:---------:|:-----------:|:-----:|:----:|:-------:|
|  1221893 |   RB  |     19    |  1663.55  | CENTREVILLE |   AL  | 2018 |    US   |
|  1221893 |   RX  |     22    |  1486.71  | CENTREVILLE |   AL  | 2018 |    US   |
|  1221893 |   RX  |     8     |   584.28  | CENTREVILLE |   AL  | 2018 |    US   |
|  1221893 |   RY  |     4     |   158.68  | CENTREVILLE |   AL  | 2018 |    US   |
|  1221894 |   RB  |     2     |   157.46  |  ALABASTER  |   AL  | 2018 |    US   |
|  1221894 |   RX  |     11    |   794.18  |  ALABASTER  |   AL  | 2018 |    US   |
|  1221894 |   RY  |     5     |   200.27  |  ALABASTER  |   AL  | 2018 |    US   |
|  1221894 |   RX  |     11    |   733.56  |  ALABASTER  |   AL  | 2018 |    US   |
|  1221894 |   RY  |     3     |   120.89  |  ALABASTER  |   AL  | 2018 |    US   |

### Analytical Framework 

<b><h4>Domain</h4></b>

Opening a New Category for the Customer. 

<b><h4>Purpose</h4></b>

1. Identify Best possible Brand Customer can be proposed given they are selling Brand A,B, and C or More. 
2. We need to consider State as 2nd variable which could impact the proposal as one of the attributes.  
	For example, I might be able to propose customer selling A from NY to start selling Brand BB, but not to as Customer Y from CL,
	as there is no enough data to prove the sales of this brand at the given state level. 
3. Recommendation would take Brand already carried by Customer to Propose the result with top 5 Brands.
4. Building a shiny app to enter the Brand and showing Result in User friendly way by applying different algorithms. 

<b><h4>Recommendation Context</h4></b>

When Customers are looking to Expand to new Category or marketing Events where customers can be targeted with new set of Category they should start selling.

<b><h4>Personalization Level </h4></b>

Based on Brand they sales today which other Brand can help them increase the revenue .

<b><h4>Interface</h4></b>

Marketing Events, Agents reaching out to customers and trying to convince  them to start the selling the new category with data we already have.

<b><h4>Whose Opinions </h4></b>

We are relaying on data from the Other customers and sales made by them in correlation with other category. 

<b><h4>Privacy and Trustworthiness </h4></b>

This data is not real, and objective is to build the system and churn it to real business use.

<b><h4>Recommendation Algorithm [Discuss within team]</h4></b>

  1. SVD / FSVD 
  2. USER to USER
  3. ITEM to ITEM 
  4.  .
  5.  .
  6.  .

### Build, Train  and Test Data {.tabset .tabset-fade .tabset-pills}

+ List all Major Variable [Discuss within team] 

| Variable Name| Desc. |
|:--------:|:-----:|
|df_ <> |For Dataframe |
| mod_<>| For Model Name |
|df_test_<>  |For Test Data|
| df_train <> |For Train Data|
|df_full <> |For full data|
|df_sample_n<>| For Sample of Size n|
|.| .|
|.| .|
|.| .|

#### Algorithm A 
#### Algorithm B
#### Algorithm C
### Evaluation of Algorithm 
### Shiny App 
### Conclusion
### Reference  

