---
title: "Project 4: Accuracy and Beyond in Restaurant Recommender Systems"
author: "Christina Valore, Juliann McEachern, & Rajwant Mishra"
date: "July 2, 2019"
output:
  prettydoc::html_pretty:
    theme: leonids 
    highlight: github
---

```{r dependencies, echo=F,comment=F,message=F,warning=F,prompt=F}
# data processing
library(tidyr); library(dplyr); library(RCurl); library(jsonlite)

#formatting
library(knitr); library(kableExtra); library(prettydoc); library(default)

# global options
## knit sizing
options(max.print="75"); opts_knit$set(width=75) 

## augment chunk output
opts_chunk$set(echo=F,cache=F, tidy=T,comment=F,message=T,warning=T) #change message/warning to F upon completion

## set table style for consistency
default(kable_styling)  <- list(bootstrap_options=c("basic"),position="center",full_width=T, font_size=10)

## working directory
try(setwd("~/Github/612-group/project-4"))
```


# Overview

The goal of this assignment is give you practice working with accuracy and other recommender system metrics. 

Tasks:

1. Compare the accuracy of at least two recommender system algorithms against your offline data. 
2. Compare and report on any change in accuracy before and after you've made the change in the algorithms. 
3. Include at least one business or user experience goal such as increased serendipity, novelty, or diversity. 
4. Textual conclusion

## Data Selection

Our data was sourced from Kaggle's [Restaurant Data with Consumer Ratings](https://www.kaggle.com/uciml/restaurant-data-with-consumer-ratings) collection, which contained several datasets pertaining to restaurants and their patrons. The cvs files are stored in our repository for replication of our work.

We subsetted the restaurant/consumer data to limit the scope of our system to one city: San Luis Potosi, Mexico. After applying transformations in the following section, we identified 90 unique users and 75 different restaurants to build our recommender systems from. 

```{r}
data1<-read.csv("./data/chefmozcuisine.csv") # cusine tags
data2<-read.csv("./data/geoplaces2.csv") # restaurant name
data3<-read.csv("./data/rating_final.csv") # user ratings
data4<-read.csv("./data/userprofile.csv") # user profile
```

## Data Tranformations

We cleaned our data using transformations and regular expression unite our user and restaurant data into a user item matrix. 

**GROUP NOTES** (DELETE LATER):

*  3 Matrices are defined. 1 has null values, 1 is imputed with the mean, 1 is imputed with 0. 
*  Dataset includes many interesting categorical features. We could try 2 user-item algorthims and a hybrid approach.
*  Lets try to do our best to minimize code chunks, add explanitory text for readers/within code chunks for us to follow.
*  Add/reference sources at the end! 

```{r}
## select attributes of interest from profile
user_profile <- data4 %>% select(userID, birth_year,budget,ambience,activity,weight)

## build user & place dataframes
user <- inner_join(data3,user_profile,by='userID') 
restaurant <- data2 %>% select(placeID, name, city,price) %>% mutate(name=tolower(gsub("[\u00ef\u00bf\u00bd\'_']", " ", name))) %>% mutate(city=tolower(city))%>%filter(grepl("slp|san|s\\.l\\.p",city)) %>% select(-city) 

## change ratings from 0-2 scale to 1-3
user$rating[user$rating==2]<-3;user$rating[user$rating==1]<-2;user$rating[user$rating==0]<-1
user$service_rating[user$service_rating==2]<-3;user$service_rating[user$service_rating==1]<-2;user$service_rating[user$service_rating==0]<-1
user$food_rating[user$food_rating==2]<-3;user$food_rating[user$food_rating==1]<-2;user$food_rating[user$food_rating==0]<-1;

## create user item matrix
ui_matrix <- inner_join(user, restaurant, by="placeID") %>% select(userID, placeID, rating) %>% spread(placeID, rating)

## build two frames for comparison - mu = mean ratings imputation; na = imputed with 0 
ui_mtx_mu <- ui_matrix; ui_mtx_na <- ui_matrix 

for(i in c(2:75)){
  ui_mtx_mu[is.na(ui_mtx_mu[,i]), i] <- round(mean(ui_mtx_mu[,i], na.rm = TRUE))
}

ui_mtx_na[is.na(ui_mtx_na)] <- 0

## preview matrices (DELETE LATER)
ui_matrix; ui_mtx_mu; ui_mtx_na

```


# Collaborative Filtering 

Business/User Goal: Serendipity

Recommend relevant items to targeted users that are different then items the user has already rated. 

## Algorithm 1
 

## Algorithm 2


## Compare 
Compare algorithms.

# Hybrid Recommender System (Juliann)

Create content and collaborative hybrid. Datasets are rich with categorical features of user & item. Reference later for [hybrid model algorithms](https://medium.com/@alfonsollanes/how-do-you-measure-and-evaluate-the-quality-of-recommendation-engines-2e91db5952af). 


Content Process: 

*  Calculate restaurant similarity using cuisine tags/price to build a vector-space representation.
*  Evaluate cosine similarity matrix 
*  Try new machine learning techniques (e.g. Naive Bayes, support vector machines, decision trees, etc) 

Pros:

*  Novelty: Relevant recommendations not based on popularity.

Cons:

*  Over-specialisation
*  May fail diversity test 

Add more later :)

# Conclusion

**As part of your textual conclusion, discuss one or more additional experiments that could be performed and/or metrics that could be evaluated only if online evaluation was possible. Also, briefly propose how you would design a reasonable online evaluation environment.**


Also: Compare content algorith outcomes to hybrid approach. 

# References:

[**Building Recommenders**:](https://buildingrecommenders.wordpress.com/2015/11/19/overview-of-recommender-algorithms-part-3/) Overview of Recommender Algorithms
