---
title: "Accuracy and Beyond in Restaurant Recommender Systems"
author: "Christina Valore, Juliann McEachern, & Rajwant Mishra"
date: "July 2, 2019"
output:
  prettydoc::html_pretty:
    theme: leonids 
    highlight: github
---

```{r dependencies, echo=F,comment=F,message=F,warning=F,prompt=F}
# data processing
library(tidyr); library(dplyr); library(RCurl); library(jsonlite); library(plyr)

#formatting
library(knitr); library(kableExtra); library(prettydoc); library(default)

#visualization
library(ggplot2)

#recommender 
library(recommenderlab); library(textmineR)

# global options
## knit sizing
options(max.print="75"); opts_knit$set(width=75) 

## augment chunk output
opts_chunk$set(echo=F,cache=F, tidy=T,comment=F,message=T,warning=T) #change message/warning to F upon completion

## set table style for consistency
default(kable_styling)  <- list(position="left",full_width=T, font_size=10)

default(scroll_box) <- list(width = "100%")

## working directory
##try(setwd("~/Github/612-group/project-4"))
```


# Overview

The goal of this assignment is give you practice working with accuracy and other recommender system metrics. 

**Assignment Tasks** (SUMMARIZE OUR PROCESS LATER):

1. Compare the accuracy of at least two recommender system algorithms against your offline data. 
2. Compare and report on any change in accuracy before and after you've made the change in the algorithms. 
3. Include at least one business or user experience goal such as increased serendipity, novelty, or diversity. 
4. Textual conclusion.

**GROUP NOTES** (DELETE LATER):

*  3 Matrices are defined. 1 has null values, 1 is imputed with the mean, 1 is imputed with 0. We do not have to use all of these. Just a starting place. 
*  Dataset includes many interesting categorical features. We could try 2 user-item algorthims and a hybrid approach.
*  Lets try to do our best to minimize code chunks, add explanitory text for readers and within code chunks for us to follow.
*  Add/reference sources at the end! 
*  Set important code chunks to `echo=T` if you want the reader to see the process.

## Data Selection

Our data was sourced from Kaggle's [Restaurant Data with Consumer Ratings](https://www.kaggle.com/uciml/restaurant-data-with-consumer-ratings) collection, which contained several datasets pertaining to restaurants and their patrons. The csv files are stored within our repository in the data folder. 

```{r}
data1<-read.csv("https://raw.githubusercontent.com/jemceach/612-group/master/project-4/data/chefmozcuisine.csv") # cusine tags
data2<-read.csv("https://raw.githubusercontent.com/jemceach/612-group/master/project-4/data/geoplaces2.csv") # restaurant name
data3<-read.csv("https://raw.githubusercontent.com/jemceach/612-group/master/project-4/data/rating_final.csv") # user ratings
data4<-read.csv("https://raw.githubusercontent.com/jemceach/612-group/master/project-4/data/userprofile.csv") # user profile
```

## Data Tranformations

We cleaned our data using transformations and regular expression unite our user and restaurant data. The output of which can be previewed below: 

```{r}
## restaurant dataframe
cuisine <- aggregate(Rcuisine ~., data1, toString)
restaurant <- data2 %>% select(placeID, name, city,price,franchise, alcohol,smoking_area) %>% mutate(name=tolower(gsub("[\u00ef\u00bf\u00bd\'_']", " ", name))) %>% mutate(city=tolower(city))
restaurant$price <- factor(restaurant$price, levels = c("low", "medium", "high")) # set factor
restaurant$city <- revalue(restaurant$city, c("cd victoria"="ciudad victoria", "cd. victoria"="ciudad victoria","victoria "="ciudad victoria","victoria"="ciudad victoria","san luis potosi "="san luis potosi","san luis potos"="san luis potosi","s.l.p"="san luis potosi","slp"="san luis potosi","s.l.p."="san luis potosi"))
restaurant$smoking_area = revalue(restaurant$smoking_area, c("none"="nonsmoking", "not permitted"="nonsmoking", "section"="smoking", "permitted"="smoking", "only at bar"="smoking"))
restaurant$price = revalue(restaurant$price, c("low"="lowprice", "medium"="mediumprice", "high" ="highprice"))
restaurant$alcohol = revalue(restaurant$alcohol, c("No_Alcohol_Served"="nonalcoholic", "Full_Bar"="fullbar", "Wine-Beer" ="winebeer"))
restaurant$city<-factor(restaurant$city, exclude = NULL)
restaurant <- inner_join(restaurant, cuisine, by="placeID")

## user dataframe: select attributes of interest from profile
user_profile <- data4 %>% select(userID, budget,activity, drink_level)
user <- inner_join(data3,user_profile,by='userID') 
user$budget <- factor(user$budget, levels = c("low", "medium", "high")) # set factor

## change ratings from 0-2 scale to 1-3
user$rating[user$rating==2]<-3;user$rating[user$rating==1]<-2;user$rating[user$rating==0]<-1
user$service_rating[user$service_rating==2]<-3;user$service_rating[user$service_rating==1]<-2;user$service_rating[user$service_rating==0]<-1
user$food_rating[user$food_rating==2]<-3;user$food_rating[user$food_rating==1]<-2;user$food_rating[user$food_rating==0]<-1;

# combine user and restaurant
data <- inner_join(user, restaurant, by="placeID") 

# subset 
data <- data %>% filter(city == "san luis potosi", activity=="student") %>% select(-city, -activity)

data %>% head() %>% kable(caption="User-Item Dataframe") %>% kable_styling() 
```


## Data Exploration

We found that 80% of our raters were students students and 76% of our restaurants were located within the Mexican city of San Luis Potosi.  As a result, we subsetted our restaurant/patron data to limit the scope of our system to this specific population. After subsetting our raw data, we identified 78 unique users and 56 restaurants to build our recommender systems from. 

The following plots help visualize the distribution of our overall ratings given by users based on their budget and the restaurant's categorized pricing. We also viewed the rating counts each restaurant received. On average, each venue received 13 user ratings. 

```{r, fig.height=2, out.width = '100%'}
ggplot(data, aes(x=rating)) + geom_histogram(bins=3, color='#000000', fill='#e4d1d1') +labs(title="Raw Ratings Distribution by Patron Budget") + facet_wrap(data$budget, nrow=1)

ggplot(data, aes(x=rating)) + geom_histogram(bins=3, color='#000000', fill='#b9b0b0') +labs(title="Raw Ratings Distribution by Restaurant Pricing") + facet_wrap(data$price, nrow=1)
```
```{r,fig.height=3, out.width = '100%'}
data %>% mutate(placeID = as.factor(placeID), rating=as.factor(rating)) %>% group_by(placeID) %>% add_tally() %>% ungroup() %>% ggplot(aes(x=reorder(placeID, -n),fill=rating))+ geom_bar(stat="count", color="#686256") + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+labs(title="Restaurant Rating Counts") + scale_fill_manual(values=c("#f0efef", "#e4d1d1", "#b9b0b0"))
```
```{r, fig.height=4.5, out.width = '100%'}
# Cuisine bar plot. May or may not include. 
data %>% select(Rcuisine, rating) %>% mutate(rating=as.factor(rating)) %>% group_by(Rcuisine) %>% add_tally() %>% ungroup %>% ggplot(aes(x=reorder(Rcuisine, -n), fill=rating)) + geom_bar(stat="count",color="#686256")+labs(title="Restaurant Cuisine Tag Counts with Ratings") + theme(axis.title.x=element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))+ scale_fill_manual(values=c("#f0efef", "#e4d1d1", "#b9b0b0"))
```


# Collaborative Filtering 

**ADD TEXT EXPLINATION OF BUSINESS/USER GOALS: Serendipity? Goal to be recommending relevant items to targeted users that are different then items the user has already rated?**

## Matrix Building 

We further transformed our data into a user-item matrix to build our recommender systems.

```{r}
# create user item matrix
ui_matrix <- data %>% select(userID, placeID, rating) %>% spread(placeID, rating)
rownames(ui_matrix)<-ui_matrix$userID # set row names to userid
ui_matrix<-ui_matrix %>% select(-userID) %>% as.matrix()# remove userid from columns
umat <- as(ui_matrix,"realRatingMatrix") # save real ratings for algo 



##### built two other frames for comparison - mu = mean ratings imputation; na = imputed with 0 
ui_mtx_mu <- ui_matrix; ui_mtx_na <- ui_matrix 

for(i in c(1:56)){
  ui_mtx_mu[is.na(ui_mtx_mu[,i]), i] <- round(mean(ui_mtx_mu[,i], na.rm = TRUE))
}

ui_mtx_na[is.na(ui_mtx_na)] <- 0

## preview matrices (DELETE UNUSED LATER)
as.data.frame.array(ui_matrix) %>% head() %>% kable(caption="User Matrix Preview (Ratings Preserved as NA)") %>% kable_styling()%>% scroll_box()
as.data.frame.array(ui_matrix) %>% head() %>% kable(caption="User Matrix Preview (NA=Mean Rating)") %>% kable_styling()%>% scroll_box()
as.data.frame.array(ui_matrix) %>% head() %>% kable(caption="User Matrix Preview (NA=0)") %>% kable_styling()%>% scroll_box()
```

## Training and Test Subsets

Our data was split into training and tests sets for model evaluation of both two recommender algorithms. We split our data using the `recommenderlab` package. 90% of data was retained for training and 10% for testing purposes.

```{r}
# evaluation method using training set of 
evalu <- evaluationScheme(umat, method="split", train=0.9, given=3, goodRating=0)
```


## Algorithm 1: User-Based (Christina)

ADD TEXT ETC. 

### Process 
Text
```{r, echo=T}
#non-normalized
ub_n <- Recommender(getData(evalu, "train"), "UBCF", 
      param=list(normalize = NULL, method="Cosine"))

#centered
ub_c <- Recommender(getData(evalu, "train"), "UBCF", 
      param=list(normalize = "center",method="Cosine"))

#Z-score normalization
ub_z <- Recommender(getData(evalu, "train"), "UBCF", 
      param=list(normalize = "Z-score",method="Cosine"))
```

### Predictions 
Text
```{r}
#predicted ratings
p1 <- predict(ub_n, getData(evalu,"known"),type="ratings")
p2 <- predict(ub_c, getData(evalu,"known"),type="ratings")
p3 <- predict(ub_z, getData(evalu,"known"),type="ratings")

#ceiling and floor values
p1@data@x[p1@data@x[] < 1] <- 1
p1@data@x[p1@data@x[] > 3] <- 3

p2@data@x[p2@data@x[] < 1] <- 1
p2@data@x[p2@data@x[] > 3] <- 3

p3@data@x[p3@data@x[] < 1] <- 1
p3@data@x[p3@data@x[] > 3] <- 3

#compare the predictions using the different normalize methods
error_UCOS <- rbind(
  ub_n = calcPredictionAccuracy(p1, getData(evalu, "unknown")),
  ub_c = calcPredictionAccuracy(p2, getData(evalu, "unknown")),
  ub_z = calcPredictionAccuracy(p3, getData(evalu, "unknown"))
)
error_UCOS %>% kable(caption="Prediction Comparisons") %>% kable_styling()
```

## Algorithm 2 (Raj)


## Compare 
Compare algorithms.

# Hybrid Recommender System (Juliann)

Create content and collaborative hybrid. Datasets are rich with categorical features of user & item. Reference later for [hybrid model algorithms](https://medium.com/@alfonsollanes/how-do-you-measure-and-evaluate-the-quality-of-recommendation-engines-2e91db5952af). 


Content Process: 

*  Calculate restaurant similarity using cuisine tags/price to build a vector-space representation.
*  Evaluate cosine similarity matrix 
*  Try new machine learning techniques (e.g. Naive Bayes, support vector machines, decision trees, etc) 

Pros:

*  Novelty: Relevant recommendations not based on popularity.

Cons:

*  Over-specialisation
*  May fail diversity test 

Add more later :)


## TFIDF

We concatenated terms describing our restaurant and ran several evaluations of term frequency-inverse document frequency (TF-IDF). We modified our initial data transformations to ensure the proper terms were evaluated properly. For example, in our initial run "beer" and "wine" were calculated at the same frequency. These terms were initial stored as beer-wine, but were separated by our `textmineR` package functions. 

We combined similar text-strings without spacing or punctuation to ensure their appropriate relevancies and concatenated these restaurant features strings as "tags" for natual language processing evaluation. The statistics of our most important words can be viewed below. 

```{r}
# select terms
terms<- data %>% select(placeID, alcohol,smoking_area,price,Rcuisine) %>% group_by(placeID) %>% distinct() %>% mutate(alcohol=as.character(alcohol), smoking_area=as.character(smoking_area), price=as.character(price), Rcuisine=as.character(Rcuisine))
terms$tags <- paste(terms$alcohol, terms$smoking_area, terms$price, terms$Rcuisine, sep=" ")

# tf-idf
dtm <- CreateDtm(doc_vec = terms$tags, doc_names = terms$placeID)
tf_mat <- TermDocFreq(dtm = dtm)

# preview output
preview <- tf_mat
rownames(preview) <- NULL 
preview %>% arrange(idf) %>% head(10) %>% kable(caption="Top TF-IDF Result") %>% kable_styling() 
```


# Conclusion

**As part of your textual conclusion, discuss one or more additional experiments that could be performed and/or metrics that could be evaluated only if online evaluation was possible. Also, briefly propose how you would design a reasonable online evaluation environment.**


Also: Compare content algorith outcomes to hybrid approach. 

# References:

[**Building Recommenders**:](https://buildingrecommenders.wordpress.com/2015/11/19/overview-of-recommender-algorithms-part-3/) Overview of Recommender Algorithms

[**recommenderlab**:](https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf)A Framework for Developing and Testing Recommendation Algorithms 
